### 2021-08-20 Arr Sched Dose1 Appts.Rmd
Update reference file. Add code to render markdown for dose 1 appointments.

---
title: "MSHS Dose 1 Vaccine Appointments"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r Reference data}
dashboard_start_date <- as.Date("8/5/21", format = "%m/%d/%y")

mandate_date <- as.Date("8/12/21", format = "%m/%d/%y")

sinai_colors <- c("#221f72", "#C4C3EF",
                  "#00AEEF", "#B3EBFF",
                  "#D80B8C", "#FBB6E2",
                  "#7F7F7F")
```

```{r Summarize and format first dose schedule data}

# Summarize first dose arrived and scheduled visits
first_dose_sch <- sched_to_date %>%
  filter(ApptDate >= dashboard_start_date &
           Dose == 1 &
           Status2 %in% c("Arr", "Sch") &
           !(Site %in% c("MSVD"))) %>%
  mutate(Timeframe = ifelse(ApptDate < mandate_date, "Before Mandate",
                            "After Mandate")) %>%
  group_by(Site, ApptDate, Status2, Timeframe) %>%
  summarize(Count = n()) %>%
  ungroup() %>%
  arrange(ApptDate, Site) %>%
  pivot_wider(names_from = c(ApptDate, Status2, Timeframe),
              values_from = Count,
              names_sep = ";") %>%
  adorn_totals(where = "row",
               name = "MSHS",
               na.rm = TRUE) %>%
  pivot_longer(cols = starts_with("2021-"),
               names_to = c("ApptDate", "Status2", "Timeframe"),
               names_sep = ";",
               values_to = "Count") %>%
  mutate(ApptDate = as.Date(ApptDate, "%Y-%m-%d"))

# Format schedule summary for use in kable
sch_summary_table <- first_dose_sch %>%
  mutate(ApptDate = format(ApptDate, "%m/%d/%y"),
         Timeframe = NULL) %>%
  pivot_wider(names_from = c(ApptDate, Status2),
              values_from = Count,
              names_sort = FALSE,
              names_sep = ";")

# Create a function for creating site level kables
first_dose_kable <- function(site) {
  site_sch_kable <- sch_summary_table %>%
    filter(Site == site)

  arr_col_start <- min(
    which(str_detect(colnames(site_sch_kable), ";Arr")))
  arr_col_end <- max(
    which(str_detect(colnames(site_sch_kable), ";Arr")))
  sch_col_start <- min(
    which(str_detect(colnames(site_sch_kable), ";Sch")))
  sch_col_end <- max(
    which(str_detect(colnames(site_sch_kable), ";Sch")))

  kable_col_names <- gsub(";.*", "", colnames(site_sch_kable))

  kable_headers <- c(" " = 1,
                   "Arrived Visits" = (arr_col_end - arr_col_start + 1),
                   "Scheduled Visits" = (sch_col_end - sch_col_start + 1))

  kable(site_sch_kable, format = "html", escape = FALSE,
        align = "c", digits = 0,
        col.names = kable_col_names,
        caption = paste0(site, " Dose 1 Summary")) %>%
    kable_styling(bootstrap_options = "hover", position = "center",
                  font_size = 11, full_width = FALSE) %>%
    column_spec(column = c(1,
                           arr_col_end,
                           sch_col_end),
                border_right = "thin solid lightgray") %>%
    column_spec(column = c(arr_col_start:arr_col_end),
                background = sinai_colors[2]) %>%
    column_spec(column = c(sch_col_start:sch_col_end),
                background = sinai_colors[4]) %>%
    add_header_above(kable_headers,
                     background = c("",
                                    sinai_colors[1],
                                    sinai_colors[3]),
                     line = FALSE,
                     color = "white") %>%
    scroll_box(width = "800px")

}

first_dose_graph <- function(site) {
  site_sch_graph <- first_dose_sch %>%
    filter(Site == site)

  ggplot(site_sch_graph,
         aes(fill = Status2,
             y = Count,
             x = ApptDate)) +
    geom_col(position = "identity") +
    labs(title = paste(site,
                       " Arrived and Scheduled Dose 1 Visits"),
         x = "Date",
         y = "Volume") +
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "bottom",
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    guides(fill = guide_legend(title = "Visit Status")) +
    geom_text(aes(x = ApptDate, y = Count, label = Count),
              # color = "white",
              # position = "identity",
              color = "black",
              position = "identity",
              vjust = -0.5, 
              # position = position_nudge(y = 0.5),
              size = 3) +
    scale_fill_manual(values = c(sinai_colors[1], sinai_colors[3]),
                      breaks = unique(site_sch_graph$Status2),
                      labels = c("Arrived", "Scheduled")) +
    scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
    scale_x_date(date_labels = "%m/%d",
                 breaks = seq.Date(from = min(site_sch_graph$ApptDate),
                                   to = max(site_sch_graph$ApptDate),
                                   by = 2),
                 expand = c(0, 1)) +
  geom_vline(xintercept = mandate_date,
             color = "#7F7F7F",
             linetype ="dashed",
             size = .75) +
  geom_text(aes(x = mandate_date, y = max(site_sch_graph$Count, na.rm = TRUE)*1.15),
            label = "Vaccine Mandate Date",
            color = "#7F7F7F",
            size = 4,
            hjust = 0)
}

```

## {.tabset}

### MSHS
```{r Create table and graph for MSHS}
first_dose_kable(site = "MSHS")
asis_output("<h4>")
first_dose_graph(site = "MSHS")
```

### MSBI
```{r Create table and graph for MSBI}
first_dose_kable(site = "MSBI")
asis_output("<h4>")
first_dose_graph(site = "MSBI")
```

### MSB
```{r Create table and graph for MSB}
first_dose_kable(site = "MSB")
asis_output("<h4>")
first_dose_graph(site = "MSB")
```

### MSH
```{r Create table and graph for MSH}
first_dose_kable(site = "MSH")
asis_output("<h4>")
first_dose_graph(site = "MSH")
```

### MSM
```{r Create table and graph for MSM}
first_dose_kable(site = "MSM")
asis_output("<h4>")
first_dose_graph(site = "MSM")
```

### MSQ
```{r Create table and graph for MSQ}
first_dose_kable(site = "MSQ")
asis_output("<h4>")
first_dose_graph(site = "MSQ")
```

### MSW
```{r Create table and graph for MSW}
first_dose_kable(site = "MSW")
asis_output("<h4>")
first_dose_graph(site = "MSW")
```

## {-}