### 2021-08-25 Empl_Arrival_Tracking.Rmd
Create code and script for daily tracking of employee vaccinations.

---
output: html_document
---

### MSHS COVID-19 Employee Vaccination Tracking
#### As of `r format(Sys.Date(), "%m/%d/%y")`
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r Create table to track employee arrivals}
empl_arr_start_date <- as.Date("8/20/21", format = "%m/%d/%y")

empl_arr_date_range <- seq.Date(from = empl_arr_start_date,
                                to = today - 1,
                                by = 1)
empl_vacc_sites <- c("MSH", "MSQ", "MSBI", "MSB", "MSM", "MSW")

empl_vacc_doses <- 1:3

empl_vacc_sites_dates <- data.frame(
  "Site" = sort(rep(empl_vacc_sites, length(empl_arr_date_range))),
  "ApptDate" = rep(empl_arr_date_range, length(empl_vacc_sites)))

empl_vacc_sites_doses <- data.frame(
  "Site" = sort(rep(empl_vacc_sites, length(empl_vacc_doses))),
  "Dose" = rep(empl_vacc_doses, length(empl_vacc_sites)))

empl_vacc_templ <- left_join(empl_vacc_sites_dates,
                             empl_vacc_sites_doses,
                             by = c("Site" = "Site"))

empl_arr_df <- employee_arr %>%
  filter(PatientType == "Employee" &
           ApptDate >= employee_arr_start_date) %>%
  mutate(PatientType = NULL)

empl_arr_df <- left_join(empl_vacc_templ,
                         empl_arr_df,
                         by = c("Site" = "Site",
                                "ApptDate" = "ApptDate",
                                "Dose" = "Dose"))

empl_arr_df_test <- empl_arr_df %>%
  mutate(Count = replace_na(Count, 0))  %>%
  pivot_wider(names_from = c(Dose, ApptDate),
              names_sep = "_",
              values_from = "Count",
              names_sort = FALSE,
              names_prefix = "Dose") %>%
  adorn_totals(where = "row",
               na.rm = TRUE,
               name = "MSHS") %>%
  pivot_longer(cols = contains("Dose"),
               names_to = c("Dose", "ApptDate"),
               names_sep = "_",
               values_to = "Count") %>%
  mutate(Dose = str_replace(Dose, "Dose", "Dose "),
         Site = factor(Site, levels = c(empl_vacc_sites, "MSHS"), ordered = TRUE),
         ApptDate = as.Date(ApptDate, format = "%Y-%m-%d")) %>%
  arrange(Dose, ApptDate, Site) %>%
  relocate(Site, .after = Dose) %>%
  mutate(ApptDate = format(ApptDate, "%m/%d/%y")) %>%
  pivot_wider(names_from = c(Dose, Site),
              names_sep = "_",
              values_from = Count) %>%
  adorn_totals(where = "row",
               na.rm = TRUE,
               name = "Total") %>%
  pivot_longer(cols = contains("Dose"),
               names_to = c("Dose", "Site"),
               names_sep = "_",
               values_to = "Count") %>%
  pivot_wider(names_from = ApptDate,
              values_from = Count,
              names_sort = FALSE)

empl_vacc_tables <- function(dose) {
  data <- empl_arr_df_test %>%
    filter(Dose == paste("Dose", dose)) %>%
    mutate(Dose = NULL)

  kable(data, format = "html", escape = FALSE,
        align = "c", digits = 0,
        caption = paste("Dose", dose, "Employee Vaccine Arrivals")) %>%
    kable_styling(bootstrap_options = "hover", position = "center",
                  font_size = 11, full_width = FALSE) %>%
    column_spec(column = c(1,
                           ncol(data) - 1,
                           ncol(data)),
                border_right = "thin solid lightgray") %>%
    column_spec(column = ncol(data),
                bold = TRUE,
                italic = TRUE) %>%
    add_header_above(c(" " = 1, "Date" = ncol(data) - 1),
                     background = c("",
                                    sinai_colors[1]),
                     line = FALSE,
                     color = "white") %>%
    row_spec(row = nrow(data),
             bold = TRUE,
             italic = TRUE,
             hline_after = TRUE)
}
```

```{r Dose 1}
empl_vacc_tables(dose = 1)

```
###

```{r Dose 2}
empl_vacc_tables(dose = 2)
```
###

```{r Dose 3}
empl_vacc_tables(dose = 3)
```

<h4><bold>Data Sources:</bold></h4>
Analysis based on Epic schedule data for arrived appointments through prior day. Employees identified as those patients who answer "yes" when asked if s/he is a Mount Sinai employee.