### 2021-11-18 Adult_Peds_Weekly_Arrivals_Sched_Tracking.Rmd
Create new script for daily ambulatory schedule tracking that includes peds vaccinations.

---
title: "MSHS COVID-19 Vaccinations"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r Install and load packages}
# Code for updating data analysis for afternoon huddle

#Install and load necessary packages --------------------
#install.packages("readraw_dfl")
#install.packages("writeraw_dfl")
#install.packages("ggplot2")
#install.packages("lubridate")
#install.packages("dplyr")
#install.packages("reshape2")
#install.packages("svDialogs")
#install.packages("stringr")
#install.packages("formattable")
#install.packages("ggpubr")
#install.packages("zipcodeR")
#install.packages("tidyr")
#install.packages("janitor")
#install.packages("knitr")
#install.packages("rmarkdown")
#install.packages("kableExtra")

#Analysis for weekend discharge tracking
library(readxl)
library(writexl)
library(ggplot2)
library(lubridate)
library(dplyr)
library(reshape2)
library(svDialogs)
library(stringr)
library(formattable)
library(scales)
library(ggpubr)
library(reshape2)
library(knitr)
library(kableExtra)
library(rmarkdown)
library(zipcodeR)
library(tidyr)
library(janitor)
library(knitr)
library(rmarkdown)
library(kableExtra)
```

```{r Import reference data, set variables, and import raw data, include = FALSE}

# Set working directory and select raw data ----------------------------
rm(list = ls())

# Determine path for working directory
if ("Presidents" %in% list.files("J://")) {
  user_directory <- paste0("J:/Presidents/HSPI-PM/",
                           "Operations Analytics and Optimization/Projects/",
                           "System Operations/COVID Vaccine")
} else {
  user_directory <- paste0("J:/deans/Presidents/HSPI-PM/",
                           "Operations Analytics and Optimization/Projects/",
                           "System Operations/COVID Vaccine")
}

# Set path for file selection
user_path <- paste0(user_directory, "\\*.*")

# Determine whether or not to update an existing repo
initial_run <- FALSE
update_repo <- TRUE

# Import reference data for site and pod mappings
site_mappings <- read_excel(paste0(user_directory, "/ScheduleData/",
                                   "Automation Ref 2021-11-03.xlsx"),
                            sheet = "Site Mappings")
pod_mappings <- read_excel(paste0(user_directory, "/ScheduleData/",
                                  "Automation Ref 2021-11-03.xlsx"),
                           sheet = "Pod Mappings Simple")

# Store today's date
today <- Sys.Date()
yesterday <- today - 1

# Site order
all_sites <- c("MSB",
               "MSBI",
               "MSH",
               "MSM",
               "MSQ",
               "MSW",
               "MSVD",
               "Network LI",
               "MSSN")

peds_sites <- c("MSH", "MSBI")

# Pod type order
pod_type <- c("Employee", "Patient", "All")

# Vaccine manufacturers
mfg <- c("Pfizer", "Moderna", "J&J")

# Vaccine types
vax_types <- c("Adult", "Peds")

# NY zip codes
ny_zips <- search_state("NY")

# Create dataframe with start dates for adult and peds vaccination
vacc_start_df <- data.frame("VaxType" = c("Adult", "Peds"),
                            "StartDate" = c(as.Date("12/15/20",
                                                    format = "%m/%d/%y"),
                                            as.Date("11/04/21",
                                                    format = "%m/%d/%y")))

vacc_start_df <- vacc_start_df %>%
  mutate(
    # Determine the Sunday of the first week of vaccination
    FirstSun = StartDate - (wday(StartDate) - 1),
    # Determine this week number
    ThisWeek = as.numeric(floor((today - FirstSun) / 7) + 1))

# Create dataframe with dates and weeks since vaccines started
all_dates <- data.frame("Date" = seq.Date(min(vacc_start_df$StartDate),
                                          today + 7,
                                          by = 1
                                          )
                        )

all_dates <- all_dates %>%
  mutate(
    # Determine vaccine week for adult vaccinations
    Adult_VaccWeek = as.numeric(
      floor((Date -
               vacc_start_df$FirstSun[
                 which(vacc_start_df$VaxType == "Adult")]) / 7) + 1),
    # Determine vaccine week for pediatric vaccinations
    Peds_VaccWeek = ifelse(
      as.numeric(
        floor((Date -
                 vacc_start_df$FirstSun[
                 which(vacc_start_df$VaxType == "Peds")]) / 7) + 1) < 1, NA,
      as.numeric(
      floor((Date -
               vacc_start_df$FirstSun[
                 which(vacc_start_df$VaxType == "Peds")]) / 7) + 1)
    )
  ) %>%
  pivot_longer(cols = contains("_VaccWeek"),
               names_to = "VaxType",
               values_to = "VaccWeek") %>%
  mutate(VaxType = str_remove(VaxType, "_VaccWeek"),
         WeekStart = Date - (wday(Date) - 1),
         WeekEnd = Date + (7 - wday(Date)),
         WeekDates = paste0(
           format(WeekStart, "%m/%d/%y"),
           "-",
           format(WeekEnd, "%m/%d/%y")
           ),
         WeekStart = NULL,
         WeekEnd = NULL
         ) %>%
  filter(!is.na(VaccWeek)) %>%
  arrange(VaxType, Date)

# Create variable with doses
doses <- c("Dose1", "Dose2", "Dose3", "AllDoses")

# Specify Sinai colors for table
sinai_colors <- c("#221f72", "#C4C3EF",
                  "#00AEEF", "#B3EBFF",
                  "#D80B8C", "#FBB6E2")

# Import raw data from Epic
if (update_repo) {
  sched_repo <- readRDS(
    choose.files(default = paste0(user_directory,
                                  "/R_Sched_AM_Repo/*.*"),
                 caption = "Select this morning's schedule repository"))

  raw_df <- read.csv(choose.files(
    default =
      paste0(user_directory, "/Epic PM Huddle Sched All Status/*.*"),
    caption = "Select this afternoon's Epic schedule"),
    stringsAsFactors = FALSE,
    check.names = FALSE)

  new_sched <- raw_df

  # Remove test patient
  new_sched <- new_sched[new_sched$Patient != "Patient, Test", ]

  # Update data classes from .csv schedule import to match repository structure
  new_sched <- new_sched %>%
    mutate(DOB = as.Date(DOB, format = "%m/%d/%Y"),
           `Made Date` = as.Date(`Made Date`, format = "%m/%d/%y"),
           Date = as.Date(Date, format = "%m/%d/%Y"),
           `Appt Time` = as.POSIXct(paste("1899-12-31 ", `Appt Time`),
                                    tz = "", format = "%Y-%m-%d %H:%M %p"),
           `ZIP Code` = substr(`ZIP Code`, 2, nchar(`ZIP Code`)))

  # Determine dates in new report
  current_dates <- sort(unique(new_sched$Date))

  # Update schedule repository by removing duplicate dates and adding data
  # from new report
  sched_repo <- sched_repo %>%
    filter(!(Date >= current_dates[1]))
  sched_repo <- rbind(sched_repo, new_sched)

  # # Save schedule
  # saveRDS(sched_repo, paste0(user_directory, "/R_Sched_PM_Huddle_Repo/",
  #                            "Epic Sched 340PM ",
  #                            format(min(sched_repo$Date), "%m%d%y"), " to ",
  #                            format(max(sched_repo$Date), "%m%d%y"),
  #                            " on ", format(Sys.time(), "%m%d%y %H%M"), ".rds"))

} else {

  sched_repo <- readRDS(
    choose.files(default = paste0(user_directory,
                                  "/R_Sched_AM_Repo/*.*"),
                 caption = "Select this morning's schedule repository"))

}

# Import MSSN data from Tableau
mssn_admin <- read_excel(
  path = choose.files(
    default = paste0(user_directory, "/MSSN Tableau Doses/*.*"),
    caption = "Select latest MSSN Tableau Administration Data"),
  col_names = TRUE,
  na = c("", NA),
  skip = 1)

mssn_admin <- mssn_admin[1:nrow(mssn_admin) - 1, ]

colnames(mssn_admin) <- c("ApptDate", "Dose1", "Dose2", "Dose3", "AllDoses")

mssn_admin <- mssn_admin %>%
  mutate(VaxType = "Adult")
```

```{r Format Epic schedule data}
# Format and analyze Epic schedule to date for dashboards ---------------------
epic_sched_to_date <- sched_repo

epic_sched_to_date <- epic_sched_to_date %>%
  mutate(
    # Determine whether appointment is for dose 1 or dose 2
    Dose = ifelse(str_detect(Type, "DOSE 1"), 1,
                  ifelse(str_detect(Type, "DOSE 2"), 2,
                         ifelse(str_detect(Type, "DOSE 3"), 3, NA))),
    # Determine appointment date, year, month, etc.
    ApptDate = date(Date),
    ApptYear = year(Date),
    ApptMonth = month(Date),
    ApptDay = day(Date),
    ApptWeek = week(Date),
    # Clean up department if it is duplicated in that column
    Department = ifelse(str_detect(Department, ","),
                        substr(Department, 1,
                               str_locate(Department, ",") - 1),
                        Department),
    # Group arrived, completed, and checked out appts as arrived
    Status = ifelse(`Appt Status` %in% c("Arrived", "Comp", "Checked Out"),
                    "Arr", `Appt Status`),
    # Update appointment status: classify any appts from prior days that are
    # still in scheduled as No Shows
    Status2 = ifelse(ApptDate < (today) & Status == "Sch", "No Show", Status),
    # Determine whether the vaccine is an adult or pediatric dose
    VaxType = ifelse(Department %in% c("1468 MADISON PEDIATRIC VACCINE POD") |
                       `Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
                                                  "DOSE 2 PEDIATRIC [1324685]"),
                     "Peds", "Adult"),
    # Determine manufacturer based on immunization and vaccine type fields
    Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
      ifelse(Status2 != "Arr", NA,
             # Assume any visits without immunization record are Pfizer
             ifelse(is.na(Immunizations) |
                      VaxType %in% c("Peds"), "Pfizer",
                    # Any immunizations with Moderna in text classified as
                    # Moderna, otherwise Pfizer
                    ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
                           # Any immunizations with Johnson and Johnson are
                           # classified as J&J
                           ifelse(str_detect(Department, "VISITING DOCS") |
                                    str_detect(Immunizations,
                                               "Johnson and Johnson"),
                                  "J&J", "Pfizer")))),
    # Determine week number of year
    WeekNum = format(ApptDate, "%U"),
    # Determine appointment day of week
    DOW = weekdays(ApptDate),
    # Determine if patient's zip code is in NYS
    NYZip = substr(`ZIP Code`, 1, 5) %in% ny_zips$zipcode
  )

# Crosswalk sites
epic_sched_to_date <- left_join(epic_sched_to_date, site_mappings,
                                by = c("Department" = "Department"))

# Manually update discrepancies in site/mfg/dose
epic_sched_to_date <- epic_sched_to_date %>%
  mutate(# Roll up any Moderna administered at MSB for YWCA to MSH
    Site = ifelse(Site %in% c("MSB") & Mfg %in% c("Moderna"), "MSH", Site),
    # Manually correct any J&J doses erroneously documented as dose 2 and change to dose 1
    Dose = ifelse(Mfg %in% c("J&J") & Dose %in% c(2), 1, Dose))

# Crosswalk pod type (employee vs patient)
epic_sched_to_date <- left_join(epic_sched_to_date,
                                pod_mappings[, c("Provider", "Pod Type")],
                                by = c("Provider/Resource" = "Provider"))

# Assume any pods without a mapping are patient pods
epic_sched_to_date[is.na(epic_sched_to_date$`Pod Type`), "Pod Type"] <- "Patient"

# Determine vaccine week based on appointment date
epic_sched_to_date <- left_join(epic_sched_to_date,
                                all_dates,
                                by = c("ApptDate" = "Date",
                                       "VaxType" = "VaxType"))

# Summarize schedule data for Epic sites by key stratification
epic_sched_summary <- epic_sched_to_date %>%
  filter(Status2 %in% c("Arr", "Sch") &
           ApptDate <= today + 7) %>%
  group_by(Site,
           VaxType,
           Dose,
           VaccWeek,
           WeekDates,
           ApptDate,
           Status2) %>%
  summarize(Count = n(),
            .groups = "keep") %>%
  pivot_wider(names_from = Dose,
              names_prefix = "Dose",
              values_from = Count) %>%
  mutate(AllDoses = sum(Dose1, Dose2, Dose3, na.rm = TRUE)) %>%
  pivot_longer(cols = c("Dose1", "Dose2", "Dose3", "AllDoses"),
               names_to = "Dose",
               values_to = "Count") %>%
  mutate(Count = replace_na(Count, 0)) %>%
  filter(!(VaxType %in% c("Peds") & Dose %in% c("Dose3"))) %>%
  ungroup()

```

```{r Format MSSN Tableau data}
# Melt MSSN data to format similar to Epic sites
mssn_admin_to_date <- mssn_admin %>%
  mutate(Site = "MSSN",
         Status2 = "Arr",
         ApptDate = as.Date(ApptDate, format = "%B %d, %Y")) %>%
  pivot_longer(cols = c("Dose1", "Dose2", "Dose3", "AllDoses"),
               names_to = "Dose",
               values_to = "Count") %>%
  mutate(Count = replace_na(Count, 0))

mssn_admin_to_date <- left_join(mssn_admin_to_date,
                                all_dates,
                                by = c("ApptDate" = "Date",
                                       "VaxType" = "VaxType"))

mssn_admin_to_date <- mssn_admin_to_date[, colnames(epic_sched_summary)]
```

```{r Combine data across all sites}
# Bind MSSN data with Epic sites data
daily_sched_summary <- rbind(epic_sched_summary, mssn_admin_to_date)

```

```{r Create and combine tables with vaccine administration and schedule data}
# Create a template of all sites, vaccine dates, and doses ---------
# Repeat dose type
doses_templ <- data.frame(
  "Dose" = rep(doses,
               length(all_sites)))

doses_templ <- doses_templ %>%
  arrange(Dose)

# Repeat vaccine types
vax_templ <- data.frame(
  "VaxType" = rep(vax_types,
                  length(all_sites)
                  )
)

vax_templ <- vax_templ %>%
  arrange(VaxType)

# Create template of sites, doses, and vaccine types
site_dose_templ <- data.frame(
  "Site" = rep(all_sites, length(doses)),
  "Dose" = doses_templ
)

site_vax_templ <- data.frame(
  "Site" = rep(all_sites, length(vax_types)),
  "VaxType" = vax_templ
)

# Manually update site-vaccine type template based on sites that are administering peds vaccine
site_vax_templ <- site_vax_templ %>%
  filter(VaxType %in% c("Adult") |
           (VaxType %in% c("Peds") & Site %in% peds_sites))

# Combine site-dose and site-vaccine type templates
site_vax_dose_templ <- left_join(site_dose_templ,
                                 site_vax_templ,
                                 by = c("Site" = "Site"))

# Remove peds dose 3
site_vax_dose_templ <- site_vax_dose_templ %>%
  filter(!(VaxType %in% c("Peds") & Dose %in% c("Dose3")))

```

```{r Subset data and create custom function for visualizing all arrived appointments to date}
vacc_weeks <- all_dates %>%
  select(VaxType, VaccWeek, WeekDates) %>%
  distinct()

site_vax_dose_week_templ <- left_join(site_vax_dose_templ,
                                      vacc_weeks,
                                      by = c("VaxType" = "VaxType")
                                      )

# Summarize arrived visits to date
arr_visits <- daily_sched_summary %>%
  filter(Status2 %in% c("Arr")) %>%
  group_by(Site,
           VaxType,
           Dose,
           VaccWeek,
           WeekDates) %>%
  summarize(Count = sum(Count, na.rm = TRUE))

# Determine the week of the last arrived visits for adults and peds vaccinations
last_arr_visits <- arr_visits %>%
  group_by(VaxType) %>%
  summarize(LatestWeek = max(VaccWeek))

# Modify the template dataframe to only include weeks with already arrived patients
arr_site_vax_dose_week_templ <- site_vax_dose_week_templ %>%
  filter((VaxType %in% c("Adult") &
            VaccWeek <=
            last_arr_visits$LatestWeek[last_arr_visits$VaxType %in% c("Adult")]
          ) |
           (VaxType %in% c("Peds") &
              VaccWeek <=
              last_arr_visits$LatestWeek[last_arr_visits$VaxType %in% c("Peds")]
            )         
         )

# Join template with all prior weeks, vaccine types, and doses with arrived visits dataframe to ensure no weeks are missing
arr_visits <- left_join(arr_site_vax_dose_week_templ %>%
                          filter(VaccWeek <= max(arr_visits$VaccWeek)),
                        arr_visits,
                        by = c("Site" = "Site",
                               "VaxType" = "VaxType",
                               "VaccWeek" = "VaccWeek",
                               "WeekDates" = "WeekDates",
                               "Dose" = "Dose"))

arr_visits <- arr_visits %>%
  mutate(Count = replace_na(Count, 0),
         WkAndDate = paste0("Wk ", VaccWeek, ": ", WeekDates))

arr_visits_kable <- function(vax_type, dose) {
  if (!(vax_type %in% c("Adult", "Peds") |
         dose %in% c(1, 2, 3, "All"))) {
    msg <- "Invalid argument"
    stop(msg)
  } else {

    arr_visits_totals <<- arr_visits %>%
      filter(VaxType %in% vax_type) %>%
      mutate(VaxType = NULL) %>%
      arrange(Dose,
              VaccWeek,
              Site) %>%
      mutate(VaccWeek = NULL,
             WeekDates = NULL) %>%
      pivot_wider(names_from = c(Dose, WkAndDate),
                  names_sep = "__",
                  values_from = Count) %>%
      adorn_totals(where = "row",
                   na.rm = TRUE,
                   name = "MSHS") %>%
      pivot_longer(cols = contains("__", ignore.case = TRUE),
                   names_to = c("Dose", "Week"),
                   names_sep = "__",
                   values_to = "Count") %>%
      pivot_wider(names_from = c(Site, Dose),
                  names_sep = "__",
                  values_from = Count) %>%
      adorn_totals(where = "row",
                   na.rm = TRUE,
                   name = "TotalDosesAdmin") %>%
      pivot_longer(cols = contains("__", ignore.case = TRUE),
                   names_to = c("Site", "Dose"),
                   names_sep = "__",
                   values_to = "Count") %>%
      relocate(Site) %>%
      relocate(Week, .after = Dose) %>%
      pivot_wider(names_from = Week,
                  values_from = Count) %>%
      relocate(TotalDosesAdmin, .after = Dose) %>%
      mutate(Site = factor(Site, levels = c(all_sites, "MSHS"),
                           ordered = TRUE),
             Dose = factor(Dose, levels = doses,
                           ordered = TRUE)) %>%
      arrange(Site, Dose)

    dose_filter <- ifelse(dose %in% c(1:3), paste0("Dose", dose),
                          paste0(dose, "Doses"))

    dose_title <- ifelse(dose %in% c(1:3), paste("Dose", dose),
                         paste (dose, "Doses"))

    arr_visits_kable_df <<- arr_visits_totals %>%
      filter(Dose == dose_filter) %>%
      mutate(Dose = NULL) %>%
      rename(`Admin Doses to Date` = TotalDosesAdmin)

    arr_visits_kable_headers <- 
      c(" " = 2,
        "Administered Doses by Week" = ncol(arr_visits_kable_df) - 2)

    kable(arr_visits_kable_df, format = "html", escape = FALSE,
        align = "c", digits = 0,
        caption = paste0(vax_type, " ", dose_title, ": Administered Doses To Date")) %>%
    kable_styling(bootstrap_options = "hover", position = "center",
                  font_size = 11, full_width = FALSE) %>%
    column_spec(column = c(1,
                           2,
                           ncol(arr_visits_kable_df)),
                border_right = "thin solid lightgray") %>%
    column_spec(column = c(3:ncol(arr_visits_kable_df)),
                background = sinai_colors[2]) %>%
    row_spec(row = nrow(arr_visits_kable_df),
             background = "lightgrey",
             bold = T,
             italic = T,
             color = "black") %>%
    add_header_above(arr_visits_kable_headers,
                     background = c("",
                                    sinai_colors[1]),
                     line = FALSE,
                     color = "white") %>%
    scroll_box(width = "800px")
  }
}
```

```{r Subset data and create custom function for visualizing schedule over next 7 days}

sch_visits <- daily_sched_summary %>%
  filter(Status2 %in% c("Sch")) %>%
  group_by(Site,
           VaxType,
           Dose,
           ApptDate) %>%
  summarize(Count = sum(Count, na.rm = TRUE))

# Create template of today through the next 7 days for each site and dose to ensure holidays are included
today_to_7days <- seq(from = today, to = today + 7, by = 1)

today_to_7days_templ <- data.frame("ApptDate" = rep(today_to_7days,
                                          length(all_sites)))

site_7day_sched_templ <- data.frame("Site" = sort(rep(all_sites,
                                                        length(today_to_7days)
                                                        )
                                                    ),
                                      "ApptDate" = today_to_7days_templ)

site_vax_dose_7day_sched_templ <- left_join(site_vax_dose_templ,
                                            site_7day_sched_templ,
                                            by = c("Site" = "Site"))

sch_visits <- left_join(site_vax_dose_7day_sched_templ,
                          sch_visits,
                          by = c("Site" = "Site",
                                 "VaxType" = "VaxType",
                                 "Dose" = "Dose",
                                 "ApptDate" = "ApptDate"))

sch_visits_kable <- function(vax_type, dose) {
  if (!(vax_type %in% c("Adult", "Peds") |
        dose %in% c(1, 2, 3, "All"))) {
    msg <- "Invalid argument"
    stop(msg)
  } else {

    sch_visits_totals <<- sch_visits %>%
      filter(VaxType %in% vax_type) %>%
      mutate(VaxType = NULL,
             Count = replace_na(Count, 0)) %>%
      arrange(Site, Dose, ApptDate) %>%
      pivot_wider(names_from = c(Dose, ApptDate),
                  names_sep = "__",
                  values_from = Count,
                  names_sort = FALSE) %>%
      adorn_totals(where = "row",
                   na.rm = TRUE,
                   name = "MSHS") %>%
      pivot_longer(cols = contains("__", ignore.case = TRUE),
                   names_to = c("Dose", "Date"),
                   names_sep = "__",
                   values_to = "Count") %>%
      mutate(Date = format(as.Date(Date, "%Y-%m-%d"), "%m/%d")) %>%
      pivot_wider(names_from = c(Site, Dose),
                  names_sep = "__",
                  values_from = Count) %>%
      adorn_totals(where = "row",
                   na.rm = TRUE,
                   name = "Total") %>%
      pivot_longer(cols = contains("__", ignore.case = TRUE),
                   names_to = c("Site", "Dose"),
                   names_sep = "__",
                   values_to = "Count") %>%
      relocate(Site) %>%
      relocate(Dose, .after = Site) %>%
      mutate(Site = factor(Site, levels = c(all_sites, "MSHS"), ordered = TRUE),
             Dose = factor(Dose, levels = doses, ordered = TRUE)) %>%
      arrange(Site, Dose) %>%
      pivot_wider(names_from = Date,
                  values_from = Count)
    dose_filter <- ifelse(dose %in% c(1:3), paste0("Dose", dose),
                          paste0(dose, "Doses"))

    dose_title <- ifelse(dose %in% c(1:3), paste("Dose", dose),
                         paste (dose, "Doses"))

    sch_visits_kable_df <<- sch_visits_totals %>%
      filter(Dose == dose_filter) %>%
      mutate(Dose = NULL)

    sch_visits_kable_headers <-
      c(" " = 1,
        "Scheduled Appointments" = ncol(sch_visits_kable_df) - 1)

    kable(sch_visits_kable_df, format = "html", escape = FALSE,
          align = "c", digits = 0,
          caption = paste0(vax_type, " ", dose_title, ": Scheduled Appointments")) %>%
      kable_styling(bootstrap_options = "hover", position = "center",
                    font_size = 11, full_width = FALSE) %>%
      column_spec(column = c(1,
                             ncol(sch_visits_kable_df) - 1,
                             ncol(sch_visits_kable_df)),
                  border_right = "thin solid lightgray") %>%
      row_spec(row = nrow(sch_visits_kable_df),
               background = "lightgrey",
               bold = T,
               italic = T,
               color = "black") %>%
      add_header_above(sch_visits_kable_headers,
                       background = c("",
                                      sinai_colors[1]),
                       line = FALSE,
                       color = "white")
  }
}

```

#### Administered & Scheduled Vaccines as of `r format(Sys.Date(), "%m/%d/%y")` 3:40PM {.tabset}

##### <span style = "color:#221f72;font-weight:bold">Adult - All Doses</span>
```{r Subset data and create kable for all adult doses administered}

arr_visits_kable(vax_type = "Adult", dose = "All")

asis_output("<h3>  </h3>")

sch_visits_kable(vax_type = "Adult", dose = "All")

```
<p></p>

##### <span style = "color:#221f72">Adult - Dose 1</span>
<p></p>
```{r Subset data and create kable for adult first doses administered}
arr_visits_kable(vax_type = "Adult", dose = 1)

asis_output("<h3>  </h3>")

sch_visits_kable(vax_type = "Adult", dose = 1)

```

<p></p>
##### <span style = "color:#221f72">Adult - Dose 2</span>
<p></p>
```{r Subset data and create kable for adult second doses administered}
arr_visits_kable(vax_type = "Adult", dose = 2)

asis_output("<h3>  </h3>")

sch_visits_kable(vax_type = "Adult", dose = 2)

```

<p></p>
##### <span style = "color:#221f72">Adult - Dose 3</span>
<p></p>
```{r Subset data and create kable for adult third doses administered}
arr_visits_kable(vax_type = "Adult", dose = 3)

asis_output("<h3>  </h3>")

sch_visits_kable(vax_type = "Adult", dose = 3)

```

<p></p>
##### <span style = "color:#00AEEF;font-weight:bold">Peds - All Doses</span>
<p></p>
```{r Subset data and create kable for all peds doses administered}
arr_visits_kable(vax_type = "Peds", dose = "All")

asis_output("<h3>  </h3>")

sch_visits_kable(vax_type = "Peds", dose = "All")

```

<p></p>
##### <span style = "color:#00AEEF">Peds - Dose 1</span>
<p></p>
```{r Subset data and create kable for peds first doses administered}
arr_visits_kable(vax_type = "Peds", dose = 1)

asis_output("<h3>  </h3>")

sch_visits_kable(vax_type = "Peds", dose = 1)

```

<p></p>
##### <span style = "color:#00AEEF">Peds - Dose 2</span>
<p></p>
```{r Subset data and create kable for peds second doses administered}
arr_visits_kable(vax_type = "Peds", dose = 2)

asis_output("<h3>  </h3>")

sch_visits_kable(vax_type = "Peds", dose = 2)

```