### 2022-04-18 MSH Admin Dose Summary.Rmd
Simplify logic for determining adult vs pediatric vaccine.

---
title: "MSHS COVID-19 Vaccination Summary"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r Install and load packages}
# Code for creating summary of COVID vaccine administration -----------------

#Install and load necessary packages --------------------
# install.packages("readxl")
# install.packages("writexl")
# install.packages("ggplot2")
# install.packages("lubridate")
# install.packages("dplyr")
# install.packages("reshape2")
# install.packages("svDialogs")
# install.packages("stringr")
# install.packages("formattable")
# install.packages("kableExtra")
# install.packages("ggpubr")
# install.packages("zipcodeR")
# install.packages("tidyr")

#Analysis for weekend discharge tracking
library(readxl)
library(writexl)
library(ggplot2)
library(lubridate)
library(dplyr)
library(reshape2)
library(svDialogs)
library(stringr)
library(formattable)
library(scales)
library(ggpubr)
library(reshape2)
library(knitr)
library(kableExtra)
library(rmarkdown)
library(zipcodeR)
library(tidyr)
library(purrr)
library(janitor)
library(DT)
```

```{r Set directory, import reference data, and repository, include = FALSE}
# # Set working directory and select raw data ----------------------------
# rm(list = ls())
# 
# # Determine path for working directory
# if ("Presidents" %in% list.files("J://")) {
#   user_directory <- paste0("J:/Presidents/HSPI-PM/",
#                            "Operations Analytics and Optimization/Projects/",
#                            "System Operations/COVID Vaccine")
# } else {
#   user_directory <- paste0("J:/deans/Presidents/HSPI-PM/",
#                            "Operations Analytics and Optimization/Projects/",
#                            "System Operations/COVID Vaccine")
# }

# Set path for file selection
user_path <- paste0(user_directory, "\\*.*")

# # Determine whether or not to update an existing repo
# update_repo <- FALSE

# Import reference data for site and pod mappings
dept_mappings <- read_excel(paste0(
  user_directory,
  "/Weekly Dashboard Doses Administered",
  "/MSHS COVID-19 Vaccine Department Mappings 2022-04-11.xlsx"))

dept_mappings <- dept_mappings %>%
  select(-`Site (Historical)`, -Notes)

# Store today's date
today <- Sys.Date()

# Site order
all_sites <- c("MSB",
               "MSBI",
               "MSH",
               "MSM",
               "MSQ",
               "MSW",
               "MSVD",
               "Network LI")

# Vaccine types
vax_types <- c("Adult", "Peds")

# Dept grouper
dept_group <- c("Hospital POD", "MSHS Practice")
```

```{r Import existing repo and append as needed}
# Import schedule repository and determine last date it was updated
sched_repo_file <- choose.files(default = paste0(user_directory,
                                                 "/R_Sched_AM_Repo/*.*"))

sched_repo <- readRDS(sched_repo_file)
sched_repo_update_date <- date(file.info(sched_repo_file)$ctime)

if (update_repo) {

  # Determine date range of reports to add to repository
  if (sched_repo_update_date == today - 1) {
    new_report_dates <- today - 1
  } else {
    new_report_dates <- seq.Date(sched_repo_update_date, today - 1, by = 1)
  }

  new_report_dates_pattern <- format(new_report_dates, "%Y%m%d")

  # Find list of reports
  file_list_vacc_sched <- list.files(
    path = paste0(user_directory,
                  "/Epic Auto Report Sched All Status"),
    pattern = paste0("^(Covid_vaccine_all_status_ZipCode_)",
                     new_report_dates_pattern,
                     ".*.csv",
                     collapse = "|"))

  # Import reports and add a column with report name and report date
  import_new_reports <- map_df(
    file_list_vacc_sched,
    .f = function(x){
      sched_data <- read.csv(
        paste0(user_directory,
               "/Epic Auto Report Sched All Status/",
               x),
        stringsAsFactors = FALSE,
        check.names = FALSE) %>%
        mutate(Filename = x,
               ReportDate = as.Date(str_extract(x, "[0-9]{8}"),
                                    "%Y%m%d"),
               DOB = as.Date(DOB, format = "%m/%d/%Y"),
               `Made Date` = as.Date(`Made Date`, format = "%m/%d/%y"),
               Date = as.Date(Date, "%m/%d/%Y"),
               `Appt Time` = as.POSIXct(paste("1899-12-31 ", `Appt Time`),
                                        tz = "", format = "%Y-%m-%d %H:%M %p"),
               `ZIP Code` = substr(`ZIP Code`, 2, nchar(`ZIP Code`)))})

  # Filter schedule data from new reports to only keep appointments in latest reports
  sched_data_new_reports <- import_new_reports %>%
    group_by(Date) %>%
    slice_max(ReportDate) %>%
    mutate(Filename = NULL,
           ReportDate = NULL) %>%
    ungroup()

  new_reports_dates <- sort(unique(sched_data_new_reports$Date))

  # Remove vaccine appoints from any overlapping dates in repository and new reports
  sched_repo <- sched_repo %>%
    filter(!(Date >= new_reports_dates[1]))

  # Bind schedule repository and schedule from new reports
  sched_repo <- rbind(sched_repo, sched_data_new_reports)

  # Save new repository
  saveRDS(sched_repo, paste0(user_directory, "/R_Sched_AM_Repo/",
                             "Auto Epic Report Sched ",
                             format(min(sched_repo$Date), "%m%d%y"), " to ",
                             format(max(sched_repo$Date), "%m%d%y"),
                             " on ", format(Sys.time(), "%m%d%y %H%M"), ".rds"))

} else {

  sched_repo <- sched_repo

}
```

```{r Format and process schedule data}
# Format and analyze schedule to date for dashboards ---------------------------
sched_to_date <- sched_repo

sched_to_date <- sched_to_date %>%
  # rename(IsPtEmployee = `Is the patient a Mount Sinai employee`) %>%
  mutate(
    # Update timezones to EST if imported as UTC
    `Appt Time` = with_tz(`Appt Time`, tzone = "EST"),
    # Determine whether appointment is for dose 1, 2, or 3 based on scheduled visit type
    Dose = ifelse(str_detect(Type, "DOSE 1"), "Dose 1",
                  ifelse(str_detect(Type, "DOSE 2"), "Dose 2",
                         ifelse(str_detect(Type, "(DOSE 3)|(BOOSTER)"),
                                "Dose 3/Boosters", NA))),
    # Determine appointment date, year, month, etc.
    ApptDate = date(Date),
    ApptYear = year(Date),
    ApptMonth = month(Date),
    ApptDay = day(Date),
    ApptWeek = week(Date),
    # Clean up department if it is duplicated in that column
    Department = ifelse(str_detect(Department, ","),
                        substr(Department, 1,
                               str_locate(Department, ",") - 1),
                        Department),
    # Group arrived, completed, and checked out appts as arrived
    Status = ifelse(`Appt Status` %in% c("Arrived", "Comp", "Checked Out"),
                    "Arr", `Appt Status`),
    # Update appointment status: classify any appts from prior days that are
    # still in scheduled as No Shows
    Status2 = ifelse(ApptDate < (today) & Status == "Sch", "No Show", 
                     # Correct any appts erroneously marked as arrived early
                     ifelse(ApptDate >= today & Status == "Arr", "Sch",
                            Status)),
    # # Determine whether the vaccine is an adult or pediatric dose
    # VaxType = ifelse(Department %in% c("1468 MADISON PEDIATRIC VACCINE POD",
    #                                    "1468 MADISON HOSP PEDS GEN") |
    #                    `Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
    #                                               "DOSE 2 PEDIATRIC [1324685]") |
    #                    Department %in% peds_school_practice_dept,
    #                  "Peds", "Adult"),
    # # Determine manufacturer based on immunization field and vaccine type fields
    # Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
    #   ifelse(Status2 != "Arr", NA,
    #          # Any blank immunizations and any peds assumed to be Pfizer
    #          ifelse(is.na(Immunizations) |
    #                   VaxType %in% c("Peds"), "Pfizer",
    #                 # Any immunizations with Moderna in text classified as Moderna
    #                 ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
    #                        # Any visiting docs or Johnson and Johnson
    #                        # immunizations classified as J&J
    #                        ifelse(str_detect(Department, "VISITING DOCS") |
    #                                 str_detect(Immunizations,
    #                                            "Johnson and Johnson"),
    #                               "J&J", "Pfizer")))),
    # Determine week number of year
    WeekNum = format(ApptDate, "%U"),
    # Determine appointment day of week
    DOW = weekdays(ApptDate)
    )

# Crosswalk sites and department grouper
sched_to_date <- left_join(sched_to_date, dept_mappings,
                           by = c("Department" = "Department"))

sched_to_date <- sched_to_date %>%
  rename(Site = `New Site`)

# Determine if there are any missing departments and notify user
unmapped_dept <- sched_to_date %>%
  filter(is.na(Site)) %>%
  select(Department) %>%
  distinct()

if(nrow(unmapped_dept) > 0) {
  stop(paste0("Check department mappings. Missing departments include ",
             paste(as.vector(unmapped_dept$Department), collapse = ", "),
             "."))
}

# Determine vaccine type (adult vs peds) administered based on department, provider, and age, as appropriate
sched_to_date <- sched_to_date %>%
  mutate(
    # Create a column for age grouper
    Age_Grouper = ifelse(str_detect(Age, "\\ y.o."), "years",
                        ifelse(str_detect(Age, "\\ m.o."), "months",
                               ifelse(str_detect(Age, "\\ days"), "days",
                                      "unknown"))),
    # Convert string with age to numeric
    Age_Numeric = 
      ifelse(Age_Grouper == "years",
                    as.numeric(str_extract(Age,
                                 # pattern = "(?<=Deceased\\ \\().+(?=\\ y.o.)")),
                                 paste("(?<=Deceased\\ \\()[0-9]+(?=\\ y.o.)",
                                       "[0-9]+(?=\\ y.o.)",
                                       sep = "|"))),
                    ifelse(Age_Grouper == "months",
                           as.numeric(str_extract(Age, "[0-9]+(?=\\ m.o.)")) / 12,
                           ifelse(Age_Grouper == "days",
                                  as.numeric(
                                    str_extract(Age, "[0-9]+(?=\\ days)")) / 365,
                                  ifelse(Age_Grouper == "unknown", 10000,
                                         NA_integer_)))),
    # Determine vaccine type based on department, provider/resource, and age if necessary
    VaxType = ifelse(
      `Vaccine Age Group` %in% "Peds" |
        # MSBI Peds POD
        `Provider/Resource` %in% c("DOSE 1 PEDIATRIC [1324684]",
                                       "DOSE 2 PEDIATRIC [1324685]") |
        # Practics that administer adult and pediatric vaccines
        (`Vaccine Age Group` %in% "Both" &
              `Setting Grouper` %in% c("School Based Practice", "MSHS Practice") &
                 Age_Numeric < 12),
      "Peds", "Adult"),
    # Determine manufacturer based on immunization field and vaccine type fields
    Mfg = #Keep manufacturer as NA if the appointment hasn't been arrived
      ifelse(Status2 != "Arr", NA,
             # Any blank immunizations and any peds assumed to be Pfizer
             ifelse(is.na(Immunizations) |
                      VaxType %in% c("Peds"), "Pfizer",
                    # Any immunizations with Moderna in text classified as Moderna
                    ifelse(str_detect(Immunizations, "Moderna"), "Moderna",
                           # Any visiting docs or Johnson and Johnson
                           # immunizations classified as J&J
                           ifelse(str_detect(Department, "VISITING DOCS") |
                                    str_detect(Immunizations,
                                               "Johnson and Johnson"),
                                  "J&J", "Pfizer")))),
    Site =
      # Roll up any Moderna administered at MSB for YWCA to MSH
      ifelse(Site %in% c("MSB") & Mfg %in% c("Moderna"), "MSH",
             # Manually classify school practices as peds or adolescent for schools administering both
             ifelse(`Setting Grouper` %in% "School Based Practice" &
                      VaxType %in% "Peds", "Peds School Based Practice",
                    ifelse(`Setting Grouper` %in% "School Based Practice" &
                             VaxType %in% "Adult",
                           "Adolescent School Based Practice",
                           Site))),
    # Manually correct any J&J doses erroneously documented as dose 2 and change to dose 1
    Dose = ifelse(Mfg %in% c("J&J") & Dose %in% c("Dose 2"), "Dose 1", Dose),
    `Setting Grouper` = ifelse(`Setting Grouper` %in% "School Based Practice",
                               "MSHS Practice", `Setting Grouper`))

vax_admin_start_date <- min(sched_to_date$ApptDate[
  which(sched_to_date$Status2 %in% "Arr")])

vax_admin_end_date <- max(sched_to_date$ApptDate[
  which(sched_to_date$Status2 %in% "Arr")])

```

#### *Vaccines Administered from `r paste0(format(vax_admin_start_date, "%m/%d/%y"), "-", format(vax_admin_end_date, "%m/%d/%y"))`*

## {.tabset}
### All Patient Settings
```{r Create kable of all doses adminsitered to date}
# Summarize data for dashboard visualizations
summary_all_doses <- sched_to_date %>%
  filter(Status2 %in% "Arr") %>%
  group_by(`Setting Grouper`, Site, VaxType) %>%
  summarize(Count = n()) %>%
  rename(Setting = `Setting Grouper`) %>%
  arrange(Setting, VaxType, -Count) %>%
  pivot_wider(names_from = VaxType,
              values_from = Count) %>%
  ungroup() %>%
  arrange(Setting, Site) %>%
  group_split(Setting) %>%
  # split(.$Setting) %>%
    # relocate(Setting, .after = Site) %>%
    map_df(., function(x) {
      adorn_totals(x, where = "row",
                        name = unique(x$Setting),
                        fill = "MSHS Total")})

summary_all_doses <- as.data.frame(summary_all_doses)

summary_all_doses <- summary_all_doses %>%
  mutate(across(.cols = where(is.numeric),
                .fns = ~replace_na(.x, 0)))

summary_all_doses[
  which(str_detect(summary_all_doses$Site, "MSHS Total")), 2:ncol(summary_all_doses)] <- 
  lapply(summary_all_doses[
    which(str_detect(summary_all_doses$Site, "MSHS Total")), 2:ncol(summary_all_doses)], 
         function(x) {cell_spec(x, italic = TRUE, bold = TRUE)})

kable(summary_all_doses,
      format = "html",
      escape = FALSE,
      align = "l",
      digits = 0,
      caption = "MSHS Total COVID-19 Vaccine Doses Administered") %>%
  kable_styling(bootstrap_options = "hover", 
                font_size = 11,
                full_width = FALSE) %>%
  row_spec(row = 0,
           background = "#221f72",
           color = "white") %>%
  column_spec(1, bold = TRUE, italic = TRUE) %>%
  collapse_rows(columns = 1)

```

```{r Custom function for creating setting specific tables}
setting_kable <- function(setting, vax_type) {

  if(!(setting %in% unique(sched_to_date$`Setting Grouper`)) |
     (!(vax_type %in% unique(sched_to_date$VaxType)))) {
    stop(paste0("Check your setting and vaccine type."))
    }

  subset_data <- sched_to_date %>%
  filter(Status2 %in% "Arr",
         `Setting Grouper` %in% setting,
         VaxType %in% vax_type) %>%
  group_by(Site, Dose) %>%
  summarize(Count = n()) %>%
  pivot_wider(names_from = Dose,
              # names_prefix = "Dose ",
              values_from = Count) %>%
  arrange(Site) %>%
  mutate(Total = sum(across(contains("Dose")), na.rm = TRUE),
         across(.cols = where(is.numeric),
                .fns = ~replace_na(.x, 0))) %>%
  adorn_totals(where = "row",
               name = "MSHS Total",
               na.rm = TRUE)

kable(subset_data,
      format = "html",
      escape = FALSE,
      align = "l",
      digits = 0,
      caption = paste(setting, vax_type,
                      "Vaccine Doses Administered"))%>%
  kable_styling(bootstrap_options = "hover", 
                font_size = 11,
                full_width = FALSE) %>%
  row_spec(row = 0,
           background = "#221f72",
           color = "white") %>%
  row_spec(row = nrow(subset_data),
           bold = TRUE,
           italic = TRUE)

}

```

### Hospital POD
```{r Create tables for hospital POD vaccine summary}

setting_kable(setting = "Hospital POD",
              vax_type = "Adult")

setting_kable(setting = "Hospital POD",
              vax_type = "Peds")

```

### MSHS Practices
```{r Create tables for MSHS practice vaccine summary}

setting_kable(setting = "MSHS Practice",
              vax_type = "Adult")

setting_kable(setting = "MSHS Practice",
              vax_type = "Peds")

```

### Reference

**Data Source**

Analysis based on Epic schedule data.

***

**Department Mappings**
```{r Reference tab with data sources and mappings}
dept_mappings_dt <- dept_mappings %>%
  # relocate(`Setting Grouper`) %>%
  relocate(Department, .after = `New Site`) %>%
  relocate(`Vaccine Age Group`, .after = Department) %>%
  rename(Site = `New Site`,
         Setting = `Setting Grouper`,
         `Vaccine Administered` = `Vaccine Age Group`) %>%
  mutate(Setting = ifelse(Setting %in% "School Based Practice",
                          "MSHS Practice", Setting),
         `Vaccine Administered` = ifelse(`Vaccine Administered` %in% "Both",
                                         "Adult & Peds",
                                         `Vaccine Administered`)) %>%
  arrange(Setting, Site, Department)

datatable(dept_mappings_dt,
          rownames = FALSE,
          options = list(
            pageLength = 20,
            paging = TRUE
          ))

```

<!-- ### School Based Practices -->
<!-- ```{r Create tables for school based practice vaccine summary} -->

<!-- setting_kable(setting = "School Based Practice", -->
<!--               vax_type = "Adult") -->

<!-- setting_kable(setting = "School Based Practice", -->
<!--               vax_type = "Peds") -->

<!-- ``` -->